---
title: "Housing Production - Overview"
author: "Brook Frye"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
source("~/affor-hous-pipe_2023/code/rs_units.R")
source("~/affor-hous-pipe_2023/code/housing_prod_general.R")
source("~/affor-hous-pipe_2023/code/afford_housing_income_levels.R")
```

#### The number of rentstabilized units over time has been steadily declining. In 2008, there were ~892,700, and in 2021, there were ~802,500. These data come from [SOA tax bills](https://a836-pts-access.nyc.gov/care/search/commonsearch.aspx?mode=persprop)

```{r}

rs_time
rs_bbl
```


#### Residential housing project completions and permits, over time by type of permit owner. Project completions have generally decreased for individual permit holders the last few years, while corporation held permits have been increasing. The permit data, which will be predictive of project completions in the future, shows a large increasing gap between individuals and corporations, with permits held by corporations increasing at a fast rate. These data come from 
[NYC DCP housing database](https://www.nyc.gov/site/planning/data-maps/open-data/dwn-housing-database.page)

```{r}
p_completions
```


```{r}
q_permits
```

#### Residential project permits by owner, 2022

```{r}
map22

```

#### Affordable housing by level of affordability, starting in 2022. These data come from HPD's [Affordable Housing Production by Project](https://data.cityofnewyork.us/Housing-Development/Affordable-Housing-Production-by-Project/hq68-rnsi)

```{r}
prop_plot
```


```{r}
m

```

##### code on [github](https://github.com/NewYorkCityCouncil/afford-hous-prod-hear_9-2023) soon.